---
title: "STA Climate Change &ndash; Cycling"
output:
#  html_document:
#    toc: true
#    toc_depth: 3
#    toc_float: true
    
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: bootstrap
    logo: sta_main_logo.png
    social: menu
 #   source_code: embed
    
---

<style>
  table {
    border-bottom: 1px solid #333;
  }
</style>

```{r global-options, include = FALSE, cache = FALSE}

knitr::opts_chunk$set(out.width = '100%', out.height = '100%', fig.align = 'left',
                      echo = FALSE, warning = FALSE, message = FALSE)

# fig.dim = c(12, 5),

```

```{r libraries_and_utilities}

source("base/common-traffic_counters.r")
source("base/common_interaction.r")

## the following libraries called in sourced base files
#library(tidyverse)
#library(lubridate)
##library(ggtext)
#library(DBI)

#library(plotly)
#library(crosstalk)

library(knitr)
library(kableExtra)
library(here)


# utilities & c.

knitr::read_chunk("base/cycle_counters_interactive.r")

```

```{r sqa_data_extraction, cache.lazy = FALSE}

#source("base/sta_climate_change_sql_extract.r", local = knitr::knit_global())
load(here("data/cycling_datasets.Rdata"))  #interim - querying and parsing from scratch
source("base/sta_climate_change_data_parsing.r", local = knitr::knit_global())
```


Overview 
=======================================================================

Row {data-height=45}
-------------------------------------
   
### Active travel data across Scotland from `r format(start_date, "%b %Y") ` to `r format(end_date, "%b %Y") `.


Column {data-height=800}
-----------------------------------------------------------------------

### Cyclists &ndash; Total Count by Month &amp; Year
```{r, bicycle_counts}
```

### Pedestrians &ndash; Total Count by Month &amp; Year
```{r, pedestrian_counts_nmf}
```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Counter Data Summaries

#### Cycling Scotland's National Monitoring Framework (NMF) Counters

```{r, summary_tables_nmf, results = 'asis'}
```

```{=html}
<p>&nbsp;<br />
Focusing on the NMF counters, at end `r format(end_date, "%b %Y") ` approximately <b>`r nmf_count_summaries["bicycle"] ` bikes</b> and <b>`r nmf_count_summaries["pedestrian"] ` pedestrians</b> had been recorded going past all counters.
Average count per day dropped in 2018 for both types of counters before starting to rise again. Looking at counts overall, <b>cycling peaked in May 2020</b>, two months after the first COVID-19 lockdown was imposed, then continued to drop after plateauing over the summer. Despite further restrictions <b>in 2021 cycling counts dropped overall</b>, even with higher counts than seen over the summer months in 2020. Counts then started to drop after the highest peak to date in Aug (2021).
<!-- This may be weather-related; icy weather at the start of 2021 and unusually cold weather and even snow in May saw very low counts, while higher counts were seen over the warmer summer months than were recorded in 2020. --><br />
 <b>Walking, overall, continued to rise in 2021.</b> But even this started to decrease in April, and by June had dropped below the same month in 2020. However, both cycling and walking started to rise again in Jul 2021, both over the 2020 figures, but then started to fall again in Aug. Walking rose in Sep before starting to fall again in Oct.
</p>
<p>&nbsp;<br />

<p>
The data now also includes council-managed counters. However, the much larger number of counters and denser network in the two largest cities, Edinburgh and Glasgow, skew the data, especially when looking at totals. We therefore compare data across location and time using predominantly average counts.
</p>

<p>&nbsp;<br />
</p>
```


#### Council Counters &amp; the John Muir Way

```{r, summary_tables_councils, results = 'asis'}
```


### Counter Locations

####
```{=html}
<p></p>
```

####

```{r, counters_installed, fig.height = 8, fig.width = 10}
```


Cycling through the Week {data-navmenu="Exploring the Data"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, filtered_data_counts_time_of_day_by_year_and_weekday}
```

### Cycling &ndash; Changes through the Week

```{r, avg_count_time_of_day_by_weekday_and_year}
```

### Change over Time &ndash; Cycling through the Week

```{r, avg_count_time_of_day_by_year_and_weekday}
```

`r rm(filtered_data) `


Cycling through the Year {data-navmenu="Exploring the Data"}
=======================================================================

> Comparing total counts for each data provider and location against averages per hour through the day... this gets more interesting looking at the data month on month, as we can see how cycling changes through the full period for which counters were installed and working. (Note - where the picture is blank no data exists for a provider for that period.)



Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, filtered_data_counts_time_of_day_by_month_and_year}
```

### Average Monthly Count

```{r, average_count_time_of_day_by_month_and_year}
```

### Monthly Totals

```{r, total_count_time_of_day_by_month_and_year}
```

`r rm(filtered_data) `


Variation across Locations {data-navmenu="Exploring the Data"}
=======================================================================

Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### By Year

```{r, average_count_by_location_and_year}
```
### By Month

```{r, average_count_by_siteID_location_and_month}
```


As the Seasons Change {data-navmenu="Exploring the Data"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Monthly Counts per Location

#### Monthly Totals - As the Seasons Change

```{r, total_count_by_location_and_month}
```

```{=html}
  <p>&nbsp;<br /></p>
```

---

#### Rainfall Patterns (mm) - Scotland

```{r, historical_weather_scotland_rainfall}
```

#### Temperature Patterns (C) - Scotland

```{r, historical_weather_scotland_temp}
```

### Changes in Counts by Year

```{r, total_count_by_location_grouped_by_year}
```



Peak Counts {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

>Counts vary from 0 to, occasionally, for cyclists, well over 20K within an hour and almost 70K in a single day. The tables below give a feel for when spikes occurred within each dataset; in East Dunbartonshire the peak hourly count occurs on the day with the peak count. Further investigation is warranted wherever an (unusual) spike is seen that does not map to an event in the location, as these sometimes simply alert to a malfunctioning counter. <br />
Any or all of day of the week, time of day and month in the year, and recent weather patterns, road type, in addition to a specific event in a location contribute to variation in numbers of especially cyclists in a location.


### Peak Counts

```{r, summary_tables_max_counts_hourly, results = 'asis'}
```

```{=html}
  <p>&nbsp;<br /></p>
  
```

```{r, summary_tables_max_counts_daily, results = 'asis'}
```


Lerwick &ndash; Change through the Seasons {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Lerwick &ndash; Change through the Seasons

#### Monthly Totals - As the Seasons Change

```{r, total_count_by_location_and_month_lerwick}
```

```{=html}
  <p>&nbsp;<br />"</p>
```
---


#### Rainfall Patterns (mm) - Lerwick, Shetland

```{r, historical_weather_lerwick_rainfall}
```

#### Temperature Patterns (C) - Lerwick, Shetland

```{r, historical_weather_lerwick_temp}
```

`r rm(counter_start, roadNames) `



About
=======================================================================


```{=html}
<p>&nbsp;<br />
<!-- A summary sheet may be downloaded &ndash; <a href="cycle_counters_info_cards.pdf">cycle_counters_info_cards.pdf</a -->

<h3>About the Visualisations</h3>

<p>
Starting with the installation of 11 bicycle and pedestrian counters by Cycling Scotland's National Monitoring Framework (NMF) in a number of counties predominantly in west, central Scotland, from `r format(start_date, "%b %Y") ` and through to `r format(end_date, "%b %Y") `. The 63rd counter was installed in Fort William in the Highlands in Aug 2021, and a further three in Edinburgh in Jun and Jul 2021. <br />
The snapshots now also include data from more dense networks of council-managed counters and some points along the John Muir Way. Some councils manage counters outwith their geographical or administrative boundaries; for instance, the City of Edinburgh Council also provides counts for Dalkeith (Midlothian). Where more than one filter is used, the charts will always group (or order) first by location (e.g., hamlet, village, town or city), then local authority, then data provider.<br />
Pedestrian count data is currently available only from NMF Counters. Summaries are shown in the overview section.
</p>

<p>
Most of the charts are interactive. Mousing over data points will display tooltips with more detail. Click/double-click on legend items to filter data sub-sets. Zoom in by drawing a rectangle around a region of interest and double-click in empty space to return to the full view. Alternatively, use the toolbar on the top, right, to interact with each chart.
</p><p>
Charts directly <a href="#variation-across-locations">comparing detail by location</a> use a <a href="https://www.investopedia.com/ask/answers/05/logvslinear.asp" title="Investopedia provide a good description of why you would use a log scale." target = "_blank">log scale (rather than linear/arithmetic)</a> because of the very large differences in counts (and counters) across locations. For instance, Edinburgh and Glasgow had ab out 40 counters each by 2018, and as at mid 2021, 57 and 75 counters respectively. Being the second and first largest (cities) by population, and Edinburgh being especially dense, directly comparing counts to all other locations dwarfs all other data, including even Stirling and Dundee &ndash; we obtain <a href="images/linear_plot_count_by_location.png" title="Why we use a log instead of a linear plot to compare counts across locations" target = "_linear_plot" >plots with most of the data sitting in the bottom, left-hand corner</a>.
</p>


<h3>Counter Locations</h3>

<p>In the overview section we list all counters &ndash; NMF and council-managed, showing installation dates and count per location. It should be noted that there is some delay between counter installation and the point from which data is recorded. Counters that came online within the last month or two may therefore not have any data reported in charts shown. Occasionally, also, a counter may go out of action temporarily and sometimes for significant periods of time.
</p>


<h3>Data Sources</h3>

<p>
Counter data may be accessed via Cycling Scotland's <a href = "https://usmart.io/org/cyclingscotland/" target = "_blank">Active Travel Open Data Portal</a>. The data feeding into this analysis is updated shortly after the end of each month, to allow for any delays in data reporting.
</p>

<p>
All data feeding into this anakysis is extracted at the level of detail of hourly counts by counter, and then aggregated as required. Data may be accessed via the API; this requires registration and request of an API key. Alternatively, a <!-- a href= "https://usmart.io/org/cyclingscotland/discovery/discovery-view-detail/7ffbe5d6-fa13-4352-829d-0bb0a58e3355" target = "_blank" -->dump may be downloaded<!-- /a --> from the portal.
</p>

<!--
City of Edinburgh Council - Daily cycling counts from automatic cycling counters
https://usmart.io/org/cyclingscotland/discovery/discovery-view-detail/b1f0bd42-e220-465e-99a3-c4f62824f21f

National Monitoring Framework - all-mode traffic survey results 2017 to 2021 - Cycling Scotland
https://usmart.io/org/cyclingscotland/discovery/discovery-view-detail/b9fdd462-1f45-4252-aedc-fbc596abcd90

NOTE TO SELF - PUT NMF ABOUT PEDESTRIAN AND ALSO EXPLANATIONS ABOUT CALCULATIONS
-->
<br />

<!-- h3>Weather Data</h3 -->
<p>
<a href = "https://www.metoffice.gov.uk/research/climate/maps-and-data/uk-and-regional-series" target = "_blank">Historical weather pattern data</a> was obtained from the Met Office.

</p>

```
