---
title: "STA Climate Change &ndash; Cycling"
output:
#  html_document:
#    toc: true
#    toc_depth: 3
#    toc_float: true
    
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: bootstrap
    social: menu
 #   source_code: embed
    
    self_contained: false
    lib_dir: libs
    
---

<style>
  table {
    border-bottom: 1px solid #333;
  }
</style>


```{css, echo = FALSE}

  .wide_image {
    overflow-x:scroll;
    white-space: nowrap;
  }

  .wide_image img {
    max-width: none;
  }


  .medium_image {
    overflow-x:scroll;
    white-space: nowrap;
  }

  .medium_image img {
    max-width: 600px;
  }

  .large_image {
    overflow:scroll;
    white-space: nowrap;
  }

  .large_image img {
    max-width: none;
    max-height: none;
  }

  .show_hide_element_box {
    #color: #fff;
    /* padding: 20px; */
    display: none;
    /* margin-top: 20px; */
  }

  .retain_default_cursor {
    cursor:default;
  }

```

```{r global-options, include = FALSE, cache = FALSE}

knitr::opts_chunk$set(out.width = '100%', out.height = '100%', fig.align = 'left',
                      echo = FALSE, warning = FALSE, message = FALSE)

# fig.dim = c(12, 5),

```

```{r libraries_and_utilities}

source("base/common-traffic_counters.r")
source("base/common_interaction.r")

## the following libraries called in sourced base files
#library(tidyverse)
#library(lubridate)
##library(ggtext)
#library(DBI)

#library(plotly)
#library(crosstalk) # no longer in use
library(RColorBrewer)

library(knitr)
library(kableExtra)


# utilities & c.

knitr::read_chunk("base/cycle_counters_interactive.r")

```

```{r sqa_data_extraction, cache.lazy = FALSE}

#source("base/sta_climate_change_sql_extract.r", local = knitr::knit_global())
load("cycling_datasets.Rdata")  #interim - querying and parsing from scratch
source("base/sta_climate_change_data_parsing.r", local = knitr::knit_global())
```


Overview 
=======================================================================

Row {data-height=45}
-------------------------------------
   
### Active travel data across Scotland from `r format(start_date, "%b %Y") ` to `r format(end_date, "%b %Y") `.


Column {data-height=620}
-----------------------------------------------------------------------

```{r, bicycle_counts_summaries}
```

### Cyclists &ndash; Total Count by Month &amp; Year
```{r, bicycle_counts}
```

### Pedestrians &ndash; Total Count by Month &amp; Year

```{r, pedestrian_counts_summaries_nmf}
```

```{r, pedestrian_counts_nmf}
```


Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Counter Data Summaries

#### Cycling Scotland's National Monitoring Framework (NMF) Counters

```{r, summary_tables_nmf_totals, results = 'asis'}
```

```{r, summary_tables_nmf_average_by_counter, results = 'asis'}
```

```{r, summary_tables_nmf_averages, results = 'asis'}
```

```{r, summary_tables_nmf_hourly_averages, results = 'asis'}
```

```{=html}
<p>&nbsp;<br />
Focusing on the NMF &dash; National Monitoring Framework &dash; counters, at end `r format(end_date, "%b %Y") ` approximately <b>`r nmf_count_summaries["bicycle"] ` bikes</b> and <b>`r nmf_count_summaries["pedestrian"] ` pedestrians</b> had been recorded going past all counters.
Average count per day dropped in 2018 for both types of counters before starting to rise again. Looking at counts overall, <b>cycling peaked in May 2020</b>, two months after the first COVID-19 lockdown was imposed, then continued to drop after plateauing over the summer. Despite further restrictions <b>in 2021 cycling counts dropped overall</b>, even with higher counts than seen over the summer months in 2020. Counts then started to drop after the highest peak to date, including also non-NMF counters, in Aug (2021) to the lowest counts recorded for 2021 in Dec, even with an increase in counters across the country.
<!-- This may be weather-related; icy weather at the start of 2021 and unusually cold weather and even snow in May saw very low counts, while higher counts were seen over the warmer summer months than were recorded in 2020. --><br />
 <b>Walking, overall, continued to rise in 2021.</b> But even this started to decrease in April, and by June had dropped below the same month in 2020. However, both cycling and walking started to rise again in Jul 2021, both over the 2020 figures, but then started to fall again in Aug. Walking rose in Sep before starting to fall again in Oct, and as for cycling, to its lowest count in 2021 in Dec.<br />
 <b>Both walking and cycling dropped in 2022, compared to 2021.</b>
</p>
<p>&nbsp;<br />

<p>
The data now also includes council-managed counters, Sustrans counters on the National Cycle Network (NCN), and counters along key points on trunk roads. Five counters along the John Muir Way, located around the west, central belt are also included in the data reported.
It should be noted that the much larger number of counters in the two largest cities, Glasgow and the particularly dense network in Edinburgh, skew the data especially when looking at totals. We therefore compare data across location and time using predominantly average and/or median counts (across locations and/or counters).
</p><p>

</p>


<p>&nbsp;<br />
</p>
```


#### Non-NMF Counters

```{r, summary_tables_councils_totals, results = 'asis'}
```

```{r, summary_tables_councils_average_by_counter, results = 'asis'}
```

```{r, summary_tables_councils_averages, results = 'asis'}
```

```{r, summary_tables_councils_hourly_averages, results = 'asis'}
```


Cycling through the Week {data-navmenu="Exploring the Data"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, filtered_data_counts_time_of_day_by_year_and_weekday}
```

### Cycling &ndash; Changes through the Week

```{r, avg_count_time_of_day_by_weekday_and_year}
```

### Change over Time &ndash; Cycling through the Week

```{r, avg_count_time_of_day_by_year_and_weekday}
```

`r rm(filtered_data) `


Cycling through the Year {data-navmenu="Exploring the Data"}
=======================================================================

> Comparing total counts for each data provider and location against averages per hour through the day... this gets more interesting looking at the data month on month, as we can see how cycling changes through the full period for which counters were installed and working. (Note - where the picture is blank no data exists for a provider for that period.)



Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

```{r, filtered_data_counts_time_of_day_by_month_and_year}
```

### Average Monthly Count

```{r, average_count_time_of_day_by_month_and_year}
```

### Monthly Totals

```{r, total_count_time_of_day_by_month_and_year}
```

`r rm(filtered_data) `


Variation across Locations {data-navmenu="Exploring the Data"}
=======================================================================

Column {.tabset .tabset-fade}
-----------------------------------------------------------------------

### By Year

```{r, median_count_by_location_and_year}
```
### By Month

```{r, median_count_by_siteID_location_and_month}
```



Peak Counts {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

>Counts vary from 0 to, occasionally, for cyclists, well over 20K within an hour and almost 70K in a single day. The tables below give a feel for when spikes occurred within each dataset; in East Dunbartonshire the peak hourly count occurs on the day with the peak count. Further investigation is warranted wherever an (unusual) spike is seen that does not map to an event in the location, as these sometimes simply alert to a malfunctioning counter. <br />
Any or all of day of the week, time of day and month in the year, and recent weather patterns, road type, in addition to a specific event in a location contribute to variation in numbers of especially cyclists in a location.


### Peak Counts

```{r, summary_tables_max_counts_hourly, results = 'asis'}
```

```{=html}
  <p>&nbsp;<br /></p>

```

```{r, summary_tables_max_counts_daily, results = 'asis'}
```


Counter Locations {data-navmenu="Data Stories"}
=======================================================================

>The timeline shows when bicycle counters came online in each location, and consistency in reporting from that point, with the most recent month at the top. Location is ordered from left to right by (highest) consistency in reporting (count greater than 0 each month). Mouse over icons in the plot to show detail or to identify points when all or no counters were reporting across each location (using the drop-down).


### Bicycle Counters &ndash; Reporting over Time

```{r, counters_reporting}
```

####

```{=html}
  <p></p>

```

<div class="retain_default_cursor">

```{r, counters_reporting_plotly, fig.height = 8, fig.width = 20}
```

</div>


What is Changing? {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### What is changing &ndash; Counters reporting or Cycling

####
```{=html}
<p>
  Comparing trends for totals (and averages) to median counts highlights locations where density (of cyclists and/or counters) varies significantly. In such cases median counts are more representative of cycling trends in a location over time, taking into account increase in reporting (and totals) due to introduction of new counters, as well as decrease due to counters not reporting data (e.g., Edinburgh in Aug 2019 and from Oct 2021, Sustrans in Jun 2019 and Perth &amp; Kinross in Dec 2021).
</p>
```

####

```{r, bicycle_counters_typical_counts_daily}
```

```{=html}
  &nbsp;

```


Cycling as the Seasons Change {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Cycle Counts by Location and Counter &ndash; As the Seasons Change

```{r, total_count_by_location_and_month}

```

####

```{=html}
  <br />&nbsp;

```

####

```{r, cycling_as_the_seasons_change}
```

```{=html}
  <br />&nbsp;

```

### Changes in Counts by Year

```{r, total_count_by_location_grouped_by_year}
```


Lerwick &ndash; Change through the Seasons {data-navmenu="Data Stories"}
=======================================================================

Row {.tabset .tabset-fade}
-----------------------------------------------------------------------

### Lerwick, Shetland &ndash; Change through the Seasons

```{r, counts_by_location_and_month_lerwick}
```

####

```{r, cycling_as_the_seasons_change_lerwick}
```

```{=html}
  <br />&nbsp;

```

`r rm(roadNames) `



About
=======================================================================

```{r, child = c("about.Rmd")}
```



```{r}
# cleanup ...

#dbDisconnect(dbConn)
#dbConn

```
