---
title: "STA Climate Change &ndash; Biannual Traffic Surveys"
output:
#  html_document:
#    toc: true
#    toc_depth: 3
#    toc_float: true
    
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: bootstrap
    social: menu
 #   source_code: embed
 
    #self_contained: false
    #lib_dir: libs
---

<style>
  table {
    border-bottom: 1px solid #333;
  }
</style>

```{r global-options, include = FALSE, cache = FALSE}

knitr::opts_chunk$set(out.width = '100%', out.height = '100%', fig.align = 'left',
                      echo = FALSE, warning = FALSE, message = FALSE)

```

```{r libraries_and_utilities}

source("base/common-traffic_counters.r")
source("base/common_interaction.r")

## the following libraries called in sourced base files
#library(tidyverse)
#library(lubridate)
##library(ggtext)
#library(DBI)

##library(plotly)
##library(crosstalk)

library(knitr)
library(kableExtra)


# utilities & c.

knitr::read_chunk("base/nmf_all_mode_traffic_surveys_interactive.r")

```

```{r surveydata_extraction, cache.lazy = FALSE}

countInterval <- "quarter_hour" # "hour" - switched to filter out query
source("base/sta_climate_change_traffic_survey_data_sql_extract.r", local = knitr::knit_global())
#load("traffic_survey_datasets.Rdata")  #interim - querying and parsing from scratch
source("base/sta_climate_change_traffic_survey_data_parsing.r", local = knitr::knit_global())

```


Overview 
=======================================================================

Row {data-height=45}
-------------------------------------
   
### Biannual Traffic Surveys from `r format(start_date, "%b %Y") ` to `r format(end_date, "%b %Y") `.


Column {data-height=650}
-----------------------------------------------------------------------

### All Transportation Modes {data-width=850}

```{r fig.height = 8, fig.width = 15}

overview_plot

```

### Cycling {.no-padding data-width=500}

```{r, sunburst_layout_bicycles, fig.width = 8}
```


Row {data-height=75}
-------------------------------------
   
### __Traffic counts by mode and type (motorised bs. active). The highest (typical daily) counts are furthest left, decreasing to the right.__

```{=html}

<table width = "100%" style = "border:none; text-align:left; color:grey" >
  <th>Motor Vehicles</th>
  <th width = "21%">Active Travel</th>
</table>

```


Column {data-height=250}
-----------------------------------------------------------------------

```{r, overview_traffic_surveys_subplots }
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[1]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[1]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[2]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[2]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[3]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[3]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[4]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[4]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[5]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[5]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[6]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[6]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[7]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[7]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[8]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[8]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[9]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[9]]
```


Daily Patterns &ndash; All Transportation  Modes {data-navmenu="Daily Patterns &ndash; All Modes"}
=======================================================================

> Typical counts (median) across all survey locations compared to the average and totals, to distinguish between growth for any transportation mode and an increase in number of locations included in the survey. Looking at both medians (more typical) and means (averages) accounts for potential skew in traffic count and transportation mode due to road type and location.

Row
-------------------------------------

### Change in Transportation Mode over Time

```{r, facet_view_survey_periods_all_modes, fig.height = 18.5, fig.width = 18}
```

### Change in Count Locations over Time

####

```{=html}

<p>&nbsp;<br />
	&nbsp;<br />
	&nbsp;<br />
	&nbsp;<br />
</p>

```

####

```{r, facet_view_survey_periods_totals_all_modes, fig.height = 5.5, fig.width = 12}
```

####

```{=html}

<p>&nbsp;<br />
	&nbsp;<br />
</p>

```

####

```{r, location_count_survey_periods_all_modes, fig.height = 5.5, fig.width = 12}
```


Daily Patterns &ndash; Cycling {data-navmenu="Daily Patterns &ndash; All Modes"}
=======================================================================

```{r, summary_bicycle_daily_patterns}
```


Row
-------------------------------------

### Typical Daily Count across Locations &ndash; Cycling

```{r, bicycle_daily_patterns_median_daily, fig.height = 6.5, fig.width = 15}
```

### Daily Average Count per Location &ndash; Cycling

```{r, bicycle_daily_patterns_average_daily, fig.height = 6.5, fig.width = 15}
```


Row
-------------------------------------

### Total Count &ndash; Cycling

```{r, bicycle_daily_patterns_count, fig.height = 6.5, fig.width = 15}
```

### No. of Locations &ndash; Cycling

```{r, bicycle_daily_patterns_count_locations, fig.height = 6.5, fig.width = 15}
```



Daily Patterns &ndash; Walking {data-navmenu="Daily Patterns &ndash; All Modes"}
=======================================================================

```{r, summary_pedestrians_daily_patterns}
```


Row
-------------------------------------

### Typical Daily Count across Locations &ndash; Walking

```{r, pedestrians_daily_patterns_median_daily, fig.height = 6.5, fig.width = 15}
```

### Daily Average Count per Location &ndash; Walking

```{r, pedestrians_daily_patterns_average_daily, fig.height = 6.5, fig.width = 15}
```


Row
-------------------------------------

### Total Count &ndash; Walking

```{r, pedestrians_daily_patterns_count, fig.height = 6.5, fig.width = 15}
```

### No. of Locations &ndash; Walking

```{r, pedestrians_daily_patterns_count_locations, fig.height = 6.5, fig.width = 15}
```


Change over Time &ndash; Active Travel {data-navmenu="Daily Patterns &ndash; All Modes"}
=======================================================================

Row
-------------------------------------

### Change over Time

```{r, growth_daily_patterns_active_travel, fig.height = 8.5, fig.width = 18}
```


Row
-------------------------------------

### Change by Season

```{r, growth_daily_patterns_seasonal_active_travel, fig.height = 8.5, fig.width = 20}
```


```{r}

 rm(max_mid_daily, summary_active_travel)
 
```



Hourly Patterns &ndash; Active Travel
=======================================================================

Row
-------------------------------------

### Cycling &ndash; per Biannual Period

```{r, facet_view_survey_periods_cyclist, fig.height = 6.5, fig.width = 25}
```

Row
-------------------------------------

### Walking &ndash; per Biannual Period

```{r, facet_view_survey_periods_pedestrian, fig.height = 6.5, fig.width = 25}
```



About
=======================================================================

```{=html}
<p>&nbsp;<br />
 The biannual survey is carried out by Scotland's National Monitoring Framework, on a Wednesday and the following Thursday normally toward the end of May and September. It tracks, on selected roads in settlements (or locations) for each of the 32 local authorities, nine modes of transport, including (passenger) cars and the two main modes of active travel &ndash; walking and cycling.
</p><br />
<p>
 Where a survey cannot be carried out on the survey dates data is typically collected on the same weekdays the following week. In a few instances there is a delay of several weeks; while this may introduce seasonality differences especially for active travel, to allow comparison over time for each location, delayed counts are included in the analysis.
</p><p>
</p>

<h4>Statistics Used</h4>

<p>
The area and sunburst charts report medians, as this provides the more meaningful, or typical count for comparing data that is often skewed by location and/or time. This allows more direct comparison between locations with large differences in number of reporting points.
</p>

<h4>Data Sources</h4>

<p>
 Data feeding into the analysis presented may be accessed via Cycling Scotland's <a href = "https://usmart.io/org/cyclingscotland/discovery/discovery-view-detail/b9fdd462-1f45-4252-aedc-fbc596abcd90" target = "_blank">Active Travel Open Data Portal</a>.
</p>
<p>&nbsp;<br />
</p>

```

Row
-------------------------------------

### Settlements with surveys completed designated survey month, cf. those partially or fully delayed per period

```{r, survey_periods, results = 'asis'}
```

```{=html}
<p>&nbsp;<br />
</p>

```
