---
title: "Greenhouse Gas Emissions in Scotland"
output:
  html_document:
    df_print: paged
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```


##### Data was obtained from the [Scottish Government report on greenhouse gas emissions ]("https://www.gov.scot/publications/scottish-greenhouse-gas-emissions-2018/") and can be downloaded from the supporting files. 

  
##### Total greenhouse gasses are the sum total of the emissions from the following: CH, CO2, HFCs, N2O, NF3, PFCs, SF6. The unit used for emissions is megatonnes of CO2, or equivalent in terms of their global warming potential for other GHGs (as calculated by IPCC described further [here]("https://www.ghgprotocol.org/sites/default/files/ghgp/Global-Warming-Potential-Values%20%28Feb%2016%202016%29_1.pdf"). Only greenhouse gas emitters have been included in this visualisation - carbon sinks in land use and forestry are not displayed. 



```{r load in packages}
library(tidyverse)
library(janitor)
library(plotly)
library(snakecase)
library(htmltools)
library(here)
```



```{r read in data}
scot_ghg <- read_csv(here("data/scottish-ghg-dataset-2018.csv")) %>%
  # fix column names
  clean_names() %>%
  # remove empty columns
  select(-starts_with("x")) %>%
  # remove the base year line
  filter(emission_year != "BaseYear") %>%
  rename(year = emission_year, 
         emissions = emissions_mt_co2e) %>%
  mutate(year = as.numeric(year)) %>%
  # select relevant columns
  select(sg_source_sector, source_name, year, pollutant, emissions)
```



```{r cleaning ghg data}
# separate out the sub categories 
scot_ghg_clean <- scot_ghg %>%
  # separate categories
  separate(col = "source_name", into = c("source_name", "source_sub_cat"), sep = " - ") %>% 
  #structure data for sunburst
  mutate(source_sub_cat = to_sentence_case(source_sub_cat),
         inner_circle = "GHG",
        # clean categories
         source_sub_cat = str_replace_all(source_sub_cat, "[Hh]gv", "HGV"),
         source_sub_cat = str_replace_all(source_sub_cat, "Lg vs", "LGVs"),
        # shorten sector names
         sg_source_sector = recode(sg_source_sector, 
                                   "Transport (excluding international aviation and shipping)" = "Domestic\nTransport",
                                   "Agriculture and Related Land Use" = "Agriculture",
                                   "Business and Industrial Process" = "Business\n&\nIndustry")) %>%
  group_by(source_name) %>%
  # amalgamate sources which only have one sub category into one, to limit the number of off shoots on sunburst
  mutate(
    source_sub_cat = case_when(
    n_distinct(source_sub_cat) == 1 ~ NA_character_,
    TRUE ~ source_sub_cat
    ),
    #wrap names for better formatting
    source_name = str_wrap(source_name, 14),
    source_sub_cat = str_wrap(source_sub_cat, 14)
    )
    
  
```



```{r preparing sunburst data}

#extract latest year and sum emissions per sector and sub categories
ghg_tree <- scot_ghg_clean %>%
    filter(year == "2018") %>%
  select(-year) %>%
  group_by(sg_source_sector, source_name, source_sub_cat) %>%
  summarise(emissions = sum(emissions)) %>%
  ungroup() %>%
  # remove carbon sinks
  filter(emissions > 0)

# create the hierarchial dataset
parents_top <- tibble(
  ids = "<b>Greenhouse Gas\nEmissions</b>",
  emissions = sum(ghg_tree$emissions),
  parents = "",
  labels = "<b>Greenhouse Gas\nEmissions</b>")


parents <- ghg_tree %>%
  group_by(sg_source_sector) %>%
  summarise(emissions = sum(emissions, na.rm = TRUE)) %>%
  mutate(parents = "<b>Greenhouse Gas\nEmissions</b>",
         labels = sg_source_sector,
         ids = paste0("<b>Greenhouse Gas\nEmissions</b> - ", sg_source_sector)) %>%
  select(-sg_source_sector)

second_level <- ghg_tree %>%
  mutate(ids = paste0("<b>Greenhouse Gas\nEmissions</b> - ", sg_source_sector, " - ", source_name),
         labels = source_name,
         parents = paste0("<b>Greenhouse Gas\nEmissions</b>", " - ", sg_source_sector)) %>%
  select(-sg_source_sector) %>%
  group_by(ids, parents, labels) %>%
  summarise(emissions = sum(emissions, na.rm = TRUE))

third_level <- ghg_tree %>%
  drop_na(source_sub_cat) %>%
  mutate(ids = paste0("<b>Greenhouse Gas\nEmissions</b> - ", sg_source_sector, " - ", source_name, " - ", source_sub_cat),
         parents = paste0("<b>Greenhouse Gas\nEmissions</b> - ", sg_source_sector, " - ", source_name)) %>%
  rename(labels = source_sub_cat) %>%
  ungroup() %>%
  select(ids, parents, labels, emissions)

# combine all levels or heirarchy
ghg_hierarchy <- bind_rows(parents_top, parents, second_level, third_level)
```

# Sources of Greenhouse Gas Emissions in Scotland in 2018
Click on a sector to expand the view

```{r warning = FALSE, message = FALSE, fig.height= 10, fig.width = 10}
library(viridis)
# set colour palette
colors <- viridis(8, direction = -1, end = 0.9)

# create sunburst
plot_ly(
  ghg_hierarchy,
  ids = ~ids,
  labels = ~ labels,
  parents = ~ parents,
  values = ~ emissions,
  type = 'sunburst',
  branchvalues = 'total',
  # orientation of text labels
  insidetextorientation='radial',
  hoverinfo = "text",
        hovertext = ~paste0("<b>",labels,"</b><br>", round(emissions, 2), " Mt")) %>%
  layout(sunburstcolorway = colors)


```



# Sources of Greenhouse Gas Emissions in Scotland from 1990 - 2018

```{r warnings = FALSE, fig.width=9, fig.height = 12}

# Calculate total greenhouse gas emissions per sector per year
ghg_sector_total <- scot_ghg_clean %>%
  group_by(year, sg_source_sector) %>%
  summarise(total_emissions = sum(emissions)) %>%
  rename(sector = sg_source_sector) %>%
  mutate( sector = recode(sector, 
                                   "Domestic\nTransport" = "Domestic Transport",
                                   "Business\n&\nIndustry" = "Business & Industry")) %>%
  filter(sector != "Forestry")

# Order the sectors by the amount they have changed over time
ghg_sector_total_plot <- ghg_sector_total %>%
  group_by(sector) %>%
  arrange(year) %>%
  mutate(annual_change = (total_emissions - lag(total_emissions))/total_emissions,
         annual_change = replace_na(annual_change, 0)) %>%
  ungroup() %>%
  mutate(sector = fct_reorder(sector, annual_change))


# Label positionings for the stacked area plot
ghg_subs <- ghg_sector_total_plot %>%
  filter(year == 1990) %>%
  ungroup() %>%
  arrange(desc(sector)) %>%
  mutate(cum = cumsum(total_emissions),
         gap = cum - lag(cum),
         gap = coalesce(gap, 0.5),
         position = cum - (gap/2))
         


#stacked area plot
  ghg_sector_total_plot %>%
  ggplot() +
  aes(x = year, y = total_emissions, group = sector, fill = sector, colour = sector) +
  geom_area() +
  theme_light() +
  scale_fill_viridis_d(option = "plasma") +
  scale_color_viridis_d(option = "plasma") +
  scale_x_continuous(breaks = seq(1990, 2018, 3)) +
  # labels
  geom_text(data = ghg_subs,
            aes(x = 1990, y = position,
                label = sector),
            colour = "black",
            hjust = 0,
            fontface = "bold") +
   theme(legend.position = "none",
         panel.grid = element_blank(),
         axis.text = element_text(face = "bold")) +
   labs(x = "Year",
        y = "Total Greenhouse Gas Emissions (Mt GWP)")
 
```




