---
title: "Home Energy"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

#Data Zone 2011 Lookup Codes Dataset:- https://www.opendata.nhs.scot/dataset/geography-codes-and-labels/resource/e92d19d4-ced7-40c8-b628-e28e4528fc41
#Domestic Energy Performance Certificates Data:- http://statistics.gov.scot/data/domestic-energy-performance-certificates
#Q3_2015 Has bad Header Row - Manually replaced


```{r}
library(shiny)
library(shinydashboard)
library(tidyverse)
library(sf)
library(here)
library(scales)
library(leaflet)
library(htmltools)
library(janitor)
library(SPARQL)
```

```{r}
#data zone lookup file
zone_lookup <- read_csv(here("raw_data/datazone_lookup.csv")) %>% clean_names()
```

```{r}
zone_lookup_clean <- zone_lookup %>% 
  select(data_zone, ca_name, ca)
```

```{r}
#Quarter 1, 2020
#endpoint <- "https://statistics.gov.scot/sparql"
#query <- "PREFIX rdfs: <http://www.w3.org/2000/01/rdf-schema#>
#SELECT ?posttown ?postcode ?co2_emissions_current_per_floor_area_kg_co2_m2_yr 
#WHERE {
#  ?obs <http://purl.org/linked-data/cube#dataSet> <http://statistics.gov.scot/data/domestic-energy-performance-certificates> .
#}"
```

# Reading in files for Quarter 1, 2, 3 & 4 for Year 2020
```{r}
q1_2020 <- read_csv(here("raw_data/ERD_EPC_data_2020_Q1_extract_0221.csv")) %>%  clean_names()
q2_2020 <- read_csv(here("raw_data/ERD_EPC_data_2020_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2020 <- read_csv(here("raw_data/ERD_EPC_data_2020_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2020 <- read_csv(here("raw_data/ERD_EPC_data_2020_Q4_extract_0221.csv")) %>%  clean_names()
```

```{r}
#removing first row (headers)
q1_2020_clean <- q1_2020[-1,] %>%
#removing unneccessary columns
    select(address1, address2, post_town, postcode, primary_energy_indicator_k_wh_m2_year, total_floor_area_m2, current_energy_efficiency_rating, current_energy_efficiency_rating_band, potential_energy_efficiency_rating, potential_energy_efficiency_rating_band,   co2_emissions_per_floor_area, current_emissions_t_co2_yr,  potential_emissions_t_co2_yr, data_zone, tenure) 
```

# Cleaning quarters 2, 3 & 4
```{r}
#cleaning quarter 2
q2_2020_clean <- q2_2020[-1,]
```

```{r}
#cleaning quarter 3
q3_2020_clean <- q3_2020[-1,]
```

```{r}
#cleaning quarter 4
q4_2020_clean <- q4_2020[-1,]
```

```{r}
#Checking for missing values across all columns
q1_2020_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

# Checking for missing values across all columns for Quarter 2
```{r}
q2_2020_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```


# Checking for missing values across all columns for Quarter 3
```{r}
q3_2020_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

# Checking for missing values across all columns for Quarter 4
```{r}
q4_2020_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

```{r}
#removing missing values from specific columns
q1_2020_clean <- q1_2020_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```

# Removing missing values from specific columns for Quarter 2
```{r}
q2_2020_clean <- q2_2020_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for Quarter 3
```{r}
q3_2020_clean <- q3_2020_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```

# Removing missing values from specific columns for Quarter 4
```{r}
q4_2020_clean <- q4_2020_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```

```{r}
q1_2020_clean$data_zone <- str_extract(q1_2020_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```

# Extracting S01 code for Quarter 2
```{r}
q2_2020_clean$data_zone <- str_extract(q2_2020_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```

# Extracting S01 code for Quarter 3
```{r}
q3_2020_clean$data_zone <- str_extract(q3_2020_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```

# Extracting S01 code for Quarter 4
```{r}
q4_2020_clean$data_zone <- str_extract(q4_2020_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```

```{r}
#converting from string to numeric
q1_2020_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2020_clean$co2_emissions_per_floor_area))

q1_2020_clean$primary_energy <- as.numeric(as.character(q1_2020_clean$primary_energy))

q1_2020_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2020_clean$current_emissions_t_co2_yr))

q1_2020_clean$potential_co2 <- as.numeric(as.character(q1_2020_clean$potential_co2))
```

#Converting from string to numeric for energy columns
```{r}
q2_2020_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2020_clean$co2_emissions_per_floor_area))

q2_2020_clean$primary_energy <- as.numeric(as.character(q2_2020_clean$primary_energy))

q2_2020_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2020_clean$current_emissions_t_co2_yr))

q2_2020_clean$potential_co2 <- as.numeric(as.character(q2_2020_clean$potential_co2))

q3_2020_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2020_clean$co2_emissions_per_floor_area))

q3_2020_clean$primary_energy <- as.numeric(as.character(q3_2020_clean$primary_energy))

q3_2020_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2020_clean$current_emissions_t_co2_yr))

q3_2020_clean$potential_co2 <- as.numeric(as.character(q3_2020_clean$potential_co2))

q4_2020_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2020_clean$co2_emissions_per_floor_area))

q4_2020_clean$primary_energy <- as.numeric(as.character(q4_2020_clean$primary_energy))

q4_2020_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2020_clean$current_emissions_t_co2_yr))

q4_2020_clean$potential_co2 <- as.numeric(as.character(q4_2020_clean$potential_co2))
```

```{r}
#adding year and quarter columns
q1_2020_clean <- q1_2020_clean %>% 
  mutate(
    year = "2020",
    quarter = "1")
```

# Adding year and quarter columns for 2, 3 & 4
```{r}
q2_2020_clean <- q2_2020_clean %>% 
  mutate(
    year = "2020",
    quarter = "2")

q3_2020_clean <- q3_2020_clean %>% 
  mutate(
    year = "2020",
    quarter = "3")

q4_2020_clean <- q4_2020_clean %>% 
  mutate(
    year = "2020",
    quarter = "4")
```


# Reading in files for Years 2012 - 2019
```{r}
q1_2019 <- read_csv(here("raw_data/ERD_EPC_data_2019_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2019 <- read_csv(here("raw_data/ERD_EPC_data_2019_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2019 <- read_csv(here("raw_data/ERD_EPC_data_2019_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2019 <- read_csv(here("raw_data/ERD_EPC_data_2019_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2018
```{r}
q1_2018 <- read_csv(here("raw_data/ERD_EPC_data_2018_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2018 <- read_csv(here("raw_data/ERD_EPC_data_2018_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2018 <- read_csv(here("raw_data/ERD_EPC_data_2018_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2018 <- read_csv(here("raw_data/ERD_EPC_data_2018_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2017
```{r}
q1_2017 <- read_csv(here("raw_data/D_EPC_data_2017_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2017 <- read_csv(here("raw_data/D_EPC_data_2017_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2017 <- read_csv(here("raw_data/D_EPC_data_2017_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2017 <- read_csv(here("raw_data/D_EPC_data_2017_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2016
```{r}
q1_2016 <- read_csv(here("raw_data/D_EPC_data_2016_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2016 <- read_csv(here("raw_data/D_EPC_data_2016_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2016 <- read_csv(here("raw_data/D_EPC_data_2016_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2016 <- read_csv(here("raw_data/D_EPC_data_2016_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2015
```{r}
q1_2015 <- read_csv(here("raw_data/D_EPC_data_2015_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2015 <- read_csv(here("raw_data/D_EPC_data_2015_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2015 <- read_csv(here("raw_data/D_EPC_data_2015_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2015 <- read_csv(here("raw_data/D_EPC_data_2015_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2014
```{r}
q1_2014 <- read_csv(here("raw_data/D_EPC_data_2014_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2014 <- read_csv(here("raw_data/D_EPC_data_2014_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2014 <- read_csv(here("raw_data/D_EPC_data_2014_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2014 <- read_csv(here("raw_data/D_EPC_data_2014_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2013
```{r}
q1_2013 <- read_csv(here("raw_data/D_EPC_data_2013_Q1_extract_0221.csv")) %>%  clean_names() 
q2_2013 <- read_csv(here("raw_data/D_EPC_data_2013_Q2_extract_0221.csv")) %>%  clean_names() 
q3_2013 <- read_csv(here("raw_data/D_EPC_data_2013_Q3_extract_0221.csv")) %>%  clean_names() 
q4_2013 <- read_csv(here("raw_data/D_EPC_data_2013_Q4_extract_0221.csv")) %>%  clean_names() 
```

#2012
```{r}
q4_2012 <- read_csv(here("raw_data/D_EPC_data_2012_Q4_extract_0221.csv")) %>%  clean_names() 
```

# Cleaning 
```{r}
# Year 2019
#removing first row
q1_2019_clean <- q1_2019[-1,]

q2_2019_clean <- q2_2019[-1,]

q3_2019_clean <- q3_2019[-1,]

q4_2019_clean <- q4_2019[-1,]
```

```{r}
# Year 2018
#removing first row
q1_2018_clean <- q1_2018[-1,]

q2_2018_clean <- q2_2018[-1,]

q3_2018_clean <- q3_2018[-1,]

q4_2018_clean <- q4_2018[-1,]
```


```{r}
# Year 2017
#removing first row
q1_2017_clean <- q1_2017[-1,]

q2_2017_clean <- q2_2017[-1,]

q3_2017_clean <- q3_2017[-1,]

q4_2017_clean <- q4_2017[-1,]
```


```{r}
# Year 2016
#removing first row
q1_2016_clean <- q1_2016[-1,]

q2_2016_clean <- q2_2016[-1,]

q3_2016_clean <- q3_2016[-1,]

q4_2016_clean <- q4_2016[-1,]
```


```{r}
# Year 2015
#removing first row
q1_2015_clean <- q1_2015[-1,]

q2_2015_clean <- q2_2015[-1,]

q3_2015_clean <- q3_2015[-1,]

q4_2015_clean <- q4_2015[-1,]
```


```{r}
# Year 2014
#removing first row
q1_2014_clean <- q1_2014[-1,]

q2_2014_clean <- q2_2014[-1,]

q3_2014_clean <- q3_2014[-1,]

q4_2014_clean <- q4_2014[-1,]
```


```{r}
# Year 2013
#removing first row
q1_2013_clean <- q1_2013[-1,]

q2_2013_clean <- q2_2013[-1,]

q3_2013_clean <- q3_2013[-1,]

q4_2013_clean <- q4_2013[-1,]
```


```{r}
# Year 2012
#removing first row
q4_2012_clean <- q4_2012[-1,]
```

# Checking for missing values across all columns for Quarter 1, 2019
```{r}
q1_2019_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```


# Checking for missing values across all columns for 2018
```{r}
q1_2018_clean %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```



# Removing missing values from specific columns for 2019
```{r}
q1_2019_clean <- q1_2019_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2019_clean <- q2_2019_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2019_clean <- q3_2019_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2019_clean <- q4_2019_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2018
```{r}
q1_2018_clean <- q1_2018_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2018_clean <- q2_2018_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2018_clean <- q3_2018_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2018_clean <- q4_2018_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2017
```{r}
q1_2017_clean <- q1_2017_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2017_clean <- q2_2017_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2017_clean <- q3_2017_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2017_clean <- q4_2017_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2016
```{r}
q1_2016_clean <- q1_2016_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2016_clean <- q2_2016_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2016_clean <- q3_2016_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2016_clean <- q4_2016_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2015
```{r}
q1_2015_clean <- q1_2015_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2015_clean <- q2_2015_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2015_clean <- q3_2015_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2015_clean <- q4_2015_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2014
```{r}
q1_2014_clean <- q1_2014_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2014_clean <- q2_2014_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2014_clean <- q3_2014_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2014_clean <- q4_2014_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2013
```{r}
q1_2013_clean <- q1_2013_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q2_2013_clean <- q2_2013_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q3_2013_clean <- q3_2013_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)

q4_2013_clean <- q4_2013_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```


# Removing missing values from specific columns for 2012
```{r}
q4_2012_clean <- q4_2012_clean %>% 
  drop_na(co2_emissions_per_floor_area, data_zone)
```

# Extracting S01 code for 2019
```{r}
q1_2019_clean$data_zone <- str_extract(q1_2019_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2019_clean$data_zone <- str_extract(q2_2019_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2019_clean$data_zone <- str_extract(q3_2019_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2019_clean$data_zone <- str_extract(q4_2019_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2018
```{r}
q1_2018_clean$data_zone <- str_extract(q1_2018_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2018_clean$data_zone <- str_extract(q2_2018_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2018_clean$data_zone <- str_extract(q3_2018_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2018_clean$data_zone <- str_extract(q4_2018_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2017
```{r}
q1_2017_clean$data_zone <- str_extract(q1_2017_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2017_clean$data_zone <- str_extract(q2_2017_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2017_clean$data_zone <- str_extract(q3_2017_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2017_clean$data_zone <- str_extract(q4_2017_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2016
```{r}
q1_2016_clean$data_zone <- str_extract(q1_2016_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2016_clean$data_zone <- str_extract(q2_2016_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2016_clean$data_zone <- str_extract(q3_2016_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2016_clean$data_zone <- str_extract(q4_2016_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2015
```{r}
q1_2015_clean$data_zone <- str_extract(q1_2015_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2015_clean$data_zone <- str_extract(q2_2015_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2015_clean$data_zone <- str_extract(q3_2015_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2015_clean$data_zone <- str_extract(q4_2015_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2014
```{r}
q1_2014_clean$data_zone <- str_extract(q1_2014_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2014_clean$data_zone <- str_extract(q2_2014_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2014_clean$data_zone <- str_extract(q3_2014_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2014_clean$data_zone <- str_extract(q4_2014_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2013
```{r}
q1_2013_clean$data_zone <- str_extract(q1_2013_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q2_2013_clean$data_zone <- str_extract(q2_2013_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q3_2013_clean$data_zone <- str_extract(q3_2013_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
q4_2013_clean$data_zone <- str_extract(q4_2013_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```


# Extracting S01 code for 2012
```{r}
q4_2012_clean$data_zone <- str_extract(q4_2012_clean$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")
```

#Converting from string to numeric for energy columns in 2019 datasets
```{r}
q1_2019_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2019_clean$co2_emissions_per_floor_area))

q1_2019_clean$primary_energy <- as.numeric(as.character(q1_2019_clean$primary_energy))

q1_2019_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2019_clean$current_emissions_t_co2_yr))

q1_2019_clean$potential_co2 <- as.numeric(as.character(q1_2019_clean$potential_co2))

q2_2019_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2019_clean$co2_emissions_per_floor_area))

q2_2019_clean$primary_energy <- as.numeric(as.character(q2_2019_clean$primary_energy))

q2_2019_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2019_clean$current_emissions_t_co2_yr))

q2_2019_clean$potential_co2 <- as.numeric(as.character(q2_2019_clean$potential_co2))

q3_2019_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2019_clean$co2_emissions_per_floor_area))

q3_2019_clean$primary_energy <- as.numeric(as.character(q3_2019_clean$primary_energy))

q3_2019_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2019_clean$current_emissions_t_co2_yr))

q3_2019_clean$potential_co2 <- as.numeric(as.character(q3_2019_clean$potential_co2))

q4_2019_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2019_clean$co2_emissions_per_floor_area))

q4_2019_clean$primary_energy <- as.numeric(as.character(q4_2019_clean$primary_energy))

q4_2019_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2019_clean$current_emissions_t_co2_yr))

q4_2019_clean$potential_co2 <- as.numeric(as.character(q4_2019_clean$potential_co2))
```


#Converting from string to numeric for energy columns in 2018 datasets
```{r}
q1_2018_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2018_clean$co2_emissions_per_floor_area))

q1_2018_clean$primary_energy <- as.numeric(as.character(q1_2018_clean$primary_energy))

q1_2018_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2018_clean$current_emissions_t_co2_yr))

q1_2018_clean$potential_co2 <- as.numeric(as.character(q1_2018_clean$potential_co2))

q2_2018_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2018_clean$co2_emissions_per_floor_area))

q2_2018_clean$primary_energy <- as.numeric(as.character(q2_2018_clean$primary_energy))

q2_2018_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2018_clean$current_emissions_t_co2_yr))

q2_2018_clean$potential_co2 <- as.numeric(as.character(q2_2018_clean$potential_co2))

q3_2018_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2018_clean$co2_emissions_per_floor_area))

q3_2018_clean$primary_energy <- as.numeric(as.character(q3_2018_clean$primary_energy))

q3_2018_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2018_clean$current_emissions_t_co2_yr))

q3_2018_clean$potential_co2 <- as.numeric(as.character(q3_2018_clean$potential_co2))

q4_2018_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2018_clean$co2_emissions_per_floor_area))

q4_2018_clean$primary_energy <- as.numeric(as.character(q4_2018_clean$primary_energy))

q4_2018_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2018_clean$current_emissions_t_co2_yr))

q4_2018_clean$potential_co2 <- as.numeric(as.character(q4_2018_clean$potential_co2))
```


#Converting from string to numeric for energy columns in 2017 datasets
```{r}
q1_2017_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2017_clean$co2_emissions_per_floor_area))

q1_2017_clean$primary_energy <- as.numeric(as.character(q1_2017_clean$primary_energy))

q1_2017_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2017_clean$current_emissions_t_co2_yr))

q1_2017_clean$potential_co2 <- as.numeric(as.character(q1_2017_clean$potential_co2))

q2_2017_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2017_clean$co2_emissions_per_floor_area))

q2_2017_clean$primary_energy <- as.numeric(as.character(q2_2017_clean$primary_energy))

q2_2017_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2017_clean$current_emissions_t_co2_yr))

q2_2017_clean$potential_co2 <- as.numeric(as.character(q2_2017_clean$potential_co2))

q3_2017_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2017_clean$co2_emissions_per_floor_area))

q3_2017_clean$primary_energy <- as.numeric(as.character(q3_2017_clean$primary_energy))

q3_2017_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2017_clean$current_emissions_t_co2_yr))

q3_2017_clean$potential_co2 <- as.numeric(as.character(q3_2017_clean$potential_co2))

q4_2017_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2017_clean$co2_emissions_per_floor_area))

q4_2017_clean$primary_energy <- as.numeric(as.character(q4_2017_clean$primary_energy))

q4_2017_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2017_clean$current_emissions_t_co2_yr))

q4_2017_clean$potential_co2 <- as.numeric(as.character(q4_2017_clean$potential_co2))
```



#Converting from string to numeric for energy columns in 2016 datasets
```{r}
q1_2016_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2016_clean$co2_emissions_per_floor_area))

q1_2016_clean$primary_energy <- as.numeric(as.character(q1_2016_clean$primary_energy))

q1_2016_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2016_clean$current_emissions_t_co2_yr))

q1_2016_clean$potential_co2 <- as.numeric(as.character(q1_2016_clean$potential_co2))

q2_2016_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2016_clean$co2_emissions_per_floor_area))

q2_2016_clean$primary_energy <- as.numeric(as.character(q2_2016_clean$primary_energy))

q2_2016_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2016_clean$current_emissions_t_co2_yr))

q2_2016_clean$potential_co2 <- as.numeric(as.character(q2_2016_clean$potential_co2))

q3_2016_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2016_clean$co2_emissions_per_floor_area))

q3_2016_clean$primary_energy <- as.numeric(as.character(q3_2016_clean$primary_energy))

q3_2016_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2016_clean$current_emissions_t_co2_yr))

q3_2016_clean$potential_co2 <- as.numeric(as.character(q3_2016_clean$potential_co2))

q4_2016_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2016_clean$co2_emissions_per_floor_area))

q4_2016_clean$primary_energy <- as.numeric(as.character(q4_2016_clean$primary_energy))

q4_2016_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2016_clean$current_emissions_t_co2_yr))

q4_2016_clean$potential_co2 <- as.numeric(as.character(q4_2016_clean$potential_co2))
```



#Converting from string to numeric for energy columns in 2015 datasets
```{r}
q1_2015_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2015_clean$co2_emissions_per_floor_area))

q1_2015_clean$primary_energy <- as.numeric(as.character(q1_2015_clean$primary_energy))

q1_2015_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2015_clean$current_emissions_t_co2_yr))

q1_2015_clean$potential_co2 <- as.numeric(as.character(q1_2015_clean$potential_co2))

q2_2015_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2015_clean$co2_emissions_per_floor_area))

q2_2015_clean$primary_energy <- as.numeric(as.character(q2_2015_clean$primary_energy))

q2_2015_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2015_clean$current_emissions_t_co2_yr))

q2_2015_clean$potential_co2 <- as.numeric(as.character(q2_2015_clean$potential_co2))

q3_2015_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2015_clean$co2_emissions_per_floor_area))

q3_2015_clean$primary_energy <- as.numeric(as.character(q3_2015_clean$primary_energy))

q3_2015_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2015_clean$current_emissions_t_co2_yr))

q3_2015_clean$potential_co2 <- as.numeric(as.character(q3_2015_clean$potential_co2))

q4_2015_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2015_clean$co2_emissions_per_floor_area))

q4_2015_clean$primary_energy <- as.numeric(as.character(q4_2015_clean$primary_energy))

q4_2015_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2015_clean$current_emissions_t_co2_yr))

q4_2015_clean$potential_co2 <- as.numeric(as.character(q4_2015_clean$potential_co2))
```



#Converting from string to numeric for energy columns in 2014 datasets
```{r}
q1_2014_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2014_clean$co2_emissions_per_floor_area))

q1_2014_clean$primary_energy <- as.numeric(as.character(q1_2014_clean$primary_energy))

q1_2014_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2014_clean$current_emissions_t_co2_yr))

q1_2014_clean$potential_co2 <- as.numeric(as.character(q1_2014_clean$potential_co2))

q2_2014_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2014_clean$co2_emissions_per_floor_area))

q2_2014_clean$primary_energy <- as.numeric(as.character(q2_2014_clean$primary_energy))

q2_2014_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2014_clean$current_emissions_t_co2_yr))

q2_2014_clean$potential_co2 <- as.numeric(as.character(q2_2014_clean$potential_co2))

q3_2014_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2014_clean$co2_emissions_per_floor_area))

q3_2014_clean$primary_energy <- as.numeric(as.character(q3_2014_clean$primary_energy))

q3_2014_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2014_clean$current_emissions_t_co2_yr))

q3_2014_clean$potential_co2 <- as.numeric(as.character(q3_2014_clean$potential_co2))

q4_2014_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2014_clean$co2_emissions_per_floor_area))

q4_2014_clean$primary_energy <- as.numeric(as.character(q4_2014_clean$primary_energy))

q4_2014_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2014_clean$current_emissions_t_co2_yr))

q4_2014_clean$potential_co2 <- as.numeric(as.character(q4_2014_clean$potential_co2))
```


#Converting from string to numeric for energy columns in 2013 datasets
```{r}
q1_2013_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q1_2013_clean$co2_emissions_per_floor_area))

q1_2013_clean$primary_energy <- as.numeric(as.character(q1_2013_clean$primary_energy))

q1_2013_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q1_2013_clean$current_emissions_t_co2_yr))

q1_2013_clean$potential_co2 <- as.numeric(as.character(q1_2013_clean$potential_co2))

q2_2013_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q2_2013_clean$co2_emissions_per_floor_area))

q2_2013_clean$primary_energy <- as.numeric(as.character(q2_2013_clean$primary_energy))

q2_2013_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q2_2013_clean$current_emissions_t_co2_yr))

q2_2013_clean$potential_co2 <- as.numeric(as.character(q2_2013_clean$potential_co2))

q3_2013_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q3_2013_clean$co2_emissions_per_floor_area))

q3_2013_clean$primary_energy <- as.numeric(as.character(q3_2013_clean$primary_energy))

q3_2013_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q3_2013_clean$current_emissions_t_co2_yr))

q3_2013_clean$potential_co2 <- as.numeric(as.character(q3_2013_clean$potential_co2))

q4_2013_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2013_clean$co2_emissions_per_floor_area))

q4_2013_clean$primary_energy <- as.numeric(as.character(q4_2013_clean$primary_energy))

q4_2013_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2013_clean$current_emissions_t_co2_yr))

q4_2013_clean$potential_co2 <- as.numeric(as.character(q4_2013_clean$potential_co2))
```

#Converting from string to numeric for energy columns in 2012 dataset

```{r}
q4_2012_clean$co2_emissions_per_floor_area <- as.numeric(as.character(q4_2012_clean$co2_emissions_per_floor_area))

q4_2012_clean$primary_energy <- as.numeric(as.character(q4_2012_clean$primary_energy))

q4_2012_clean$current_emissions_t_co2_yr <- as.numeric(as.character(q4_2012_clean$current_emissions_t_co2_yr))

q4_2012_clean$potential_co2 <- as.numeric(as.character(q4_2012_clean$potential_co2))
```




# Adding year and quarter columns for 2019
```{r}
q1_2019_clean <- q1_2019_clean %>% 
  mutate(
    year = "2019",
    quarter = "1")

q2_2019_clean <- q2_2019_clean %>% 
  mutate(
    year = "2019",
    quarter = "2")

q3_2019_clean <- q3_2019_clean %>% 
  mutate(
    year = "2019",
    quarter = "3")

q4_2019_clean <- q4_2019_clean %>% 
  mutate(
    year = "2019",
    quarter = "4")
```


# Adding year and quarter columns for 2018
```{r}
q1_2018_clean <- q1_2018_clean %>% 
  mutate(
    year = "2018",
    quarter = "1")

q2_2018_clean <- q2_2018_clean %>% 
  mutate(
    year = "2018",
    quarter = "2")

q3_2018_clean <- q3_2018_clean %>% 
  mutate(
    year = "2018",
    quarter = "3")

q4_2018_clean <- q4_2018_clean %>% 
  mutate(
    year = "2018",
    quarter = "4")
```


# Adding year and quarter columns for 2017
```{r}
q1_2017_clean <- q1_2017_clean %>% 
  mutate(
    year = "2017",
    quarter = "1")

q2_2017_clean <- q2_2017_clean %>% 
  mutate(
    year = "2017",
    quarter = "2")

q3_2017_clean <- q3_2017_clean %>% 
  mutate(
    year = "2017",
    quarter = "3")

q4_2017_clean <- q4_2017_clean %>% 
  mutate(
    year = "2017",
    quarter = "4")
```


# Adding year and quarter columns for 2016
```{r}
q1_2016_clean <- q1_2016_clean %>% 
  mutate(
    year = "2016",
    quarter = "1")

q2_2016_clean <- q2_2016_clean %>% 
  mutate(
    year = "2016",
    quarter = "2")

q3_2016_clean <- q3_2016_clean %>% 
  mutate(
    year = "2016",
    quarter = "3")

q4_2016_clean <- q4_2016_clean %>% 
  mutate(
    year = "2016",
    quarter = "4")
```


# Adding year and quarter columns for 2015
```{r}
q1_2015_clean <- q1_2015_clean %>% 
  mutate(
    year = "2015",
    quarter = "1")

q2_2015_clean <- q2_2015_clean %>% 
  mutate(
    year = "2015",
    quarter = "2")

q3_2015_clean <- q3_2015_clean %>% 
  mutate(
    year = "2015",
    quarter = "3")

q4_2015_clean <- q4_2015_clean %>% 
  mutate(
    year = "2015",
    quarter = "4")
```


# Adding year and quarter columns for 2014
```{r}
q1_2014_clean <- q1_2014_clean %>% 
  mutate(
    year = "2014",
    quarter = "1")

q2_2014_clean <- q2_2014_clean %>% 
  mutate(
    year = "2014",
    quarter = "2")

q3_2014_clean <- q3_2014_clean %>% 
  mutate(
    year = "2014",
    quarter = "3")

q4_2014_clean <- q4_2014_clean %>% 
  mutate(
    year = "2014",
    quarter = "4")
```


# Adding year and quarter columns for 2013
```{r}
q1_2013_clean <- q1_2013_clean %>% 
  mutate(
    year = "2013",
    quarter = "1")

q2_2013_clean <- q2_2013_clean %>% 
  mutate(
    year = "2013",
    quarter = "2")

q3_2013_clean <- q3_2013_clean %>% 
  mutate(
    year = "2013",
    quarter = "3")

q4_2013_clean <- q4_2013_clean %>% 
  mutate(
    year = "2013",
    quarter = "4")
```


# Adding year and quarter columns for 2012
```{r}
q4_2012_clean <- q4_2012_clean %>% 
  mutate(
    year = "2012",
    quarter = "4")
```

# Joining 2012 - 2020 data 
```{r}
home_energy <- rbind(q1_2020_clean, q2_2020_clean, q3_2020_clean, q4_2020_clean, 
                     q1_2019_clean, q2_2019_clean, q3_2019_clean, q4_2019_clean,
                     q1_2018_clean, q2_2018_clean, q3_2018_clean, q4_2018_clean, 
                     q1_2017_clean, q2_2017_clean, q3_2017_clean, q4_2017_clean,
                     q1_2015_clean, q2_2015_clean, q4_2015_clean, q3_2015_clean,
                    q1_2016_clean, q2_2016_clean, q3_2016_clean, q4_2016_clean,
                    q1_2014_clean, q2_2014_clean, q3_2014_clean, q4_2014_clean, 
                     q1_2013_clean, q2_2013_clean, q3_2013_clean, q4_2013_clean,
                     q4_2012_clean) 
                     
```

```{r}
#Hard coding to clean some town names
home_energy <- home_energy %>% 
  mutate(
    post_town = if_else(post_town == "glasgow city", "glasgow", post_town),
    post_town = if_else(post_town == "anstruther,fife", "anstruther", post_town),
    post_town = if_else(post_town == "perth &amp; kinross", "perth and kinross", post_town),
    post_town = if_else(post_town == "st. andrews", "st andrews", post_town),
    post_town = if_else(post_town == "linithgow", "linlithgow", post_town),
    post_town = if_else(post_town == "gorebeidge", "gorebridge", post_town),
    post_town = if_else(post_town == "by kyle of lochalsh, ross-shire", "kyle of lochalsh", post_town),
    post_town = if_else(post_town == "kyle", "kyle of lochalsh", post_town),
    post_town = if_else(post_town == "balfron station", "balfron", post_town),
    post_town = if_else(post_town == "\"denny \"", "denny", post_town),
    post_town = if_else(post_town == "\"glasgow \"", "glasgow", post_town),
    post_town = if_else(post_town == "\"edinburgh \"", "edinburgh", post_town)
  )
```

```{r}
#converting town names to lower case to reduce multiple instances of the same town
	home_energy$post_town = tolower(home_energy$post_town)
```

```{r}
#removing rogue white spaces from town names
home_energy$post_town = str_trim(home_energy$post_town)
```

#Joining 2019 & 2020 datasets with datazone_lookup
```{r}
home_energy_zone <- home_energy %>% 
  left_join(zone_lookup_clean, by = "data_zone")
```



```{r}
home_energy_zone %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```


```{r}
home_energy_zone <- home_energy_zone %>% 
  drop_na(data_zone)
```

```{r}
#removing unknown tenure
#home_energy_zone <- home_energy_zone %>% 
#  mutate(
#    tenure = na_if(tenure, "unknown")) %>% 
#  filter(!is.na(tenure))
```


#Writing to csv
```{r}
write_csv(home_energy_zone, (here("clean_data/home_energy.csv")))
```


