---
title: "Home Energy"
output: html_notebook
editor_options: 
  chunk_output_type: console
---

#Data Zone 2001 Lookup Codes Dataset:- https://www.opendata.nhs.scot/dataset/geography-codes-and-labels/resource/e92d19d4-ced7-40c8-b628-e28e4528fc41

#Domestic Energy Performance Certificates Data:- http://statistics.gov.scot/data/domestic-energy-performance-certificates

```{r}
library(shiny)
library(shinydashboard)
library(tidyverse)
library(sf)
library(here)
library(scales)
library(leaflet)
library(htmltools)
library(janitor)
library(SPARQL)
library(readr)
```

```{r}
#data zone lookup file
zone_lookup <- read_csv(here("raw_data/datazone_lookup.csv")) %>% clean_names()
```

```{r}
zone_lookup_clean <- zone_lookup %>% 
  select(data_zone, ca_name, ca)
```

# identify number of extract csv files in directory
```{r}
#the pattern matches the latest downloaded data file
setwd("raw_data")
fils <- list.files(pattern = "*_0721.csv")
no_of_files <- length(fils)
```

# converting files to utf-8
```{r}
for (val in fils)
{
#make sure all files are UTF-8
temp_file <- read.csv(val)
write.csv(temp_file, val, fileEncoding = "UTF-8")
file.remove(temp_file)
}
```

# Main process loop - for each quarter file
```{r}
home_energy <- data.frame()

for (i in 1:length(fils))
{
#Read in each csv file in the folder
    tmp_df <- read.csv(fils[i])
    
#removing first row (headers) create temp df, remove first row, make column names lower case
    tmp_df <- tmp_df[-1,]
    tmp_df <- tmp_df %>% 
    clean_names()
#clean problem names from the conversion of the squared character
    colnames(tmp_df)[colnames(tmp_df) == 'primary_energy_indicator_k_wh_m_a_a2_year'] <- 'primary_energy_indicator_k_wh_m2_year'
    colnames(tmp_df)[colnames(tmp_df) == 'total_floor_area_m_a_a2'] <- 'total_floor_area_m2'
    colnames(tmp_df)[colnames(tmp_df) == 'co2_emissions_current_per_floor_area_kg_co2_m_a_a2_yr'] <- 'co2_emissions_current_per_floor_area_kg_co2_m2_yr'
    
#keeping relevant columns    
    tmp_df <- tmp_df  %>%
    select(property_uprn, address1, address2, post_town, postcode, date_of_assessment, primary_energy_indicator_k_wh_m2_year, total_floor_area_m2, current_energy_efficiency_rating, current_energy_efficiency_rating_band, potential_energy_efficiency_rating, potential_energy_efficiency_rating_band, co2_emissions_current_per_floor_area_kg_co2_m2_yr, current_emissions_t_co2_yr, potential_reduction_in_emissions_t_co2_yr, data_zone, tenure, quarterly_report) %>%
    
#removing missing values from specific columns
  drop_na(co2_emissions_current_per_floor_area_kg_co2_m2_yr, data_zone)
    
#Checking for missing values across all columns
tmp_df %>% 
  summarise(across(.fns = ~sum(is.na(.x))))

# Extracting S01 code
tmp_df$data_zone <- str_extract(tmp_df$data_zone, "[Ss][0-9][0-9][0-9][0-9][0-9][0-9][0-9][0-9]")

# capturing which report each assessment comes from
tmp_df$quarterly_report = fils[i]

# Converting from string to date
tmp_df$date_of_assessment <- as.Date(tmp_df$date_of_assessment, format="%Y-%m-%d")

# extracting the date in to bits
tmp_df$year_of_assessment <- as.numeric(format(tmp_df$date_of_assessment, "%Y"))
tmp_df$month_of_assessment <- as.numeric(format(tmp_df$date_of_assessment, "%m"))
tmp_df$day_of_assessment <- as.numeric(format(tmp_df$date_of_assessment, "%d"))

#converting from string to numeric
tmp_df$co2_emissions_current_per_floor_area_kg_co2_m2_yr <- as.numeric(as.character(tmp_df$co2_emissions_current_per_floor_area_kg_co2_m2_yr))
tmp_df$primary_energy_indicator_k_wh_m2_year <- as.numeric(as.character(tmp_df$primary_energy_indicator_k_wh_m2_year))
tmp_df$current_emissions_t_co2_yr <- as.numeric(as.character(tmp_df$current_emissions_t_co2_yr))
tmp_df$potential_reduction_in_emissions_t_co2_yr <- as.numeric(as.character(tmp_df$potential_reduction_in_emissions_t_co2_yr))

#combine in to one dataframe
home_energy <- rbind.data.frame(home_energy, tmp_df)

}

```

#Hard coding to clean some town names
```{r}

home_energy <- home_energy %>% 
  mutate(
    post_town = if_else(post_town == "glasgow city", "glasgow", post_town),
    post_town = if_else(post_town == "anstruther,fife", "anstruther", post_town),
    post_town = if_else(post_town == "perth &amp; kinross", "perth and kinross", post_town),
    post_town = if_else(post_town == "st. andrews", "st andrews", post_town),
    post_town = if_else(post_town == "linithgow", "linlithgow", post_town),
    post_town = if_else(post_town == "gorebeidge", "gorebridge", post_town),
    post_town = if_else(post_town == "by kyle of lochalsh, ross-shire", "kyle of lochalsh", post_town),
    post_town = if_else(post_town == "kyle", "kyle of lochalsh", post_town),
    post_town = if_else(post_town == "balfron station", "balfron", post_town),
    post_town = if_else(post_town == "\"denny \"", "denny", post_town),
    post_town = if_else(post_town == "\"glasgow \"", "glasgow", post_town),
    post_town = if_else(post_town == "\"edinburgh \"", "edinburgh", post_town)
  )
```

#converting town names to lower case to reduce multiple instances of the same town
```{r}
	home_energy$post_town = tolower(home_energy$post_town)
```

#removing rogue white spaces from town names
```{r}
home_energy$post_town = str_trim(home_energy$post_town)
```

#Hard coding to fix some bad dates
```{r}

home_energy <- home_energy %>% 
  mutate(
    date_of_assessment = if_else(year_of_assessment == "13", paste("2013", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "14", paste("2014", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "15", paste("2015", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "16", paste("2016", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "21", paste("2021", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "213", paste("2013", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "1970", paste("2016", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "2002", paste("2021", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "2103", paste("2013", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
    date_of_assessment = if_else(year_of_assessment == "2108", paste("2018", month_of_assessment, day_of_assessment, sep = "-"), as.character(date_of_assessment)),
  )
home_energy$date_of_assessment <- as.Date(home_energy$date_of_assessment, format="%Y-%m-%d")
```

#calculating the future CO2 emmissions
```{r}
	home_energy <- home_energy %>%
  mutate(future_emmissions = (current_emissions_t_co2_yr - potential_reduction_in_emissions_t_co2_yr))
```

#Joining with datazone_lookup
```{r}
home_energy_zone <- home_energy %>% 
  left_join(zone_lookup_clean, by = "data_zone")
```

```{r}
home_energy_zone %>% 
  summarise(across(.fns = ~sum(is.na(.x))))
```

```{r}
home_energy_zone <- home_energy_zone %>% 
  drop_na(data_zone)
```

#Writing to csv
```{r}
write_csv(home_energy_zone, (here("clean_data/home_energy.csv")))
```
