---
title: "STA Climate Change &ndash; Biannual Traffic Surveys"
output:
#  html_document:
#    toc: true
#    toc_depth: 3
#    toc_float: true
    
  flexdashboard::flex_dashboard:
    orientation: rows
    theme: bootstrap
    social: menu
 #   source_code: embed
    #self_contained: false
    #lib_dir: libs
---

<style>
  table {
    border-bottom: 1px solid #333;
  }
</style>

```{r global-options, include = FALSE, cache = FALSE}

knitr::opts_chunk$set(out.width = '100%', out.height = '100%', fig.align = 'left',
                      echo = FALSE, warning = FALSE, message = FALSE)

```

```{r libraries_and_utilities}

source("base/common-traffic_counters.r")
source("base/common_interaction.r")

## the following libraries called in sourced base files
#library(tidyverse)
#library(lubridate)
##library(ggtext)
#library(DBI)

##library(plotly)
##library(crosstalk)

library(knitr)
library(kableExtra)


# utilities & c.

knitr::read_chunk("base/nmf_all_mode_traffic_surveys_interactive.r")

```

```{r sqa_data_extraction, cache.lazy = FALSE}

countInterval <- "hour"
#source("base/sta_climate_change_traffic_survey_data_sql_extract.r", local = knitr::knit_global())
load("traffic_survey_datasets.Rdata")  #interim - querying and parsing from scratch
source("base/sta_climate_change_traffic_survey_data_parsing.r", local = knitr::knit_global())

```


Overview 
=======================================================================

Row {data-height=45}
-------------------------------------
   
### Biannual Traffic Surveys from `r format(start_date, "%b %Y") ` to `r format(end_date, "%b %Y") `.


Column {data-height=650}
-----------------------------------------------------------------------

### All Transportation Modes {data-width=850}

```{r fig.height = 8, fig.width = 15}

overview_plot

```

### Cycling {.no-padding}

```{r, sunburst_layout_bicycles}
```


Row {data-height=75}
-------------------------------------
   
### __Traffic counts by mode. The highest counts are furthest left, decreasing to the right.__

```{=html}

<table width = "100%" style = "border:none; text-align:left; color:grey" >
  <th>Motor Vehicles</th>
  <th width = "21%">Active Travel</th>
</table>

```


Column {data-height=250}
-----------------------------------------------------------------------

```{r, overview_traffic_surveys_subplots }
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[1]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[1]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[2]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[2]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[3]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[3]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[4]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[4]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[5]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[5]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[6]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[6]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[7]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[7]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[8]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[8]]
```

### `r transportation_modes[levels(overview_plot_colour_scheme$TransportationMode)[9]] ` {data-width=100}

```{r, fig.height= 5}
transportation_mode_plots[[9]]
```



About
=======================================================================


```{=html}
<!-- p>&nbsp;<br />

<p>
</p -->


<h3>Data Sources</h3>

<p>
 Data feeding into the analysis presented may be accessed via Cycling Scotland's <a href = "https://usmart.io/org/cyclingscotland/discovery/discovery-view-detail/b9fdd462-1f45-4252-aedc-fbc596abcd90" target = "_blank">Active Travel Open Data Portal</a>.
</p>

```
