---
title: Scottish Tech Army Climate Dashboard
output: 
  flexdashboard::flex_dashboard:
    theme: bootstrap
    logo: sta_main_logo.png
    orientation: rows
    social: menu
    self_contained: true
    lib_dir: libs
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(janitor)
library(plotly)
library(snakecase)
library(here)
library(viridis)

```

```{r common code}
source("./STAClimateDashboard/code/r/index_ghg/index_markdown/index_base/simp_chart_plot.R") 
```

```{r read in data}
scot_ghg <- read_csv("data/scottish-ghg-dataset-2019.csv", show_col_types = FALSE) %>%
  # fix column names
  clean_names() %>%
  # remove empty columns
  select(-starts_with("x")) %>%
  # remove the base year line
  filter(emission_year != "BaseYear", ghg_emissions_mt_co2e > 0) %>%
  rename(year = emission_year, 
         emissions = ghg_emissions_mt_co2e) %>%
  mutate(year = as.numeric(year)) %>%
  # select relevant columns
  select(sg_categories, source_name, year, pollutant, emissions)
```

```{r cleaning ghg data}
# separate out the sub categories 
scot_ghg_clean <- scot_ghg %>%
  # separate categories
  separate(col = "source_name", into = c("source_name", "source_sub_cat"), sep = " - ") %>% 
  #structure data for sunburst
  mutate(source_sub_cat = to_sentence_case(source_sub_cat),
         inner_circle = "GHG",
        # clean categories
         source_sub_cat = str_replace_all(source_sub_cat, "[Hh]gv", "HGV"),
         source_sub_cat = str_replace_all(source_sub_cat, "Lg vs", "LGVs"),
        # shorten sector names
         sg_categories = recode(sg_categories, 
                                   "Transport (excluding international aviation and shipping)" = "Domestic\nTransport",
                                   "Agriculture and Related Land Use" = "Agriculture",
                                   "Business and Industrial Process" = "Business\n&\nIndustry")) %>%
  group_by(source_name) %>%
  # amalgamate sources which only have one sub category into one, to limit the number of off shoots on sunburst
  mutate(
    source_sub_cat = case_when(
    n_distinct(source_sub_cat) == 1 ~ NA_character_,
    TRUE ~ source_sub_cat
    ),
    #wrap names for better formatting
    source_name = str_wrap(source_name, 14),
    source_sub_cat = str_wrap(source_sub_cat, 14)
    )
```
```{r identify last year and GHG choice}
#first bit
last_year <- "2019"
emission_type <- "Scottish Greenhouse Gas "
```

```{r preparing sunburst data}

#extract latest year and sum emissions per sector and sub categories
ghg_tree <- scot_ghg_clean %>%
  filter(year == last_year) %>%
  select(-year) %>%
  group_by(sg_categories, source_name, source_sub_cat) %>%
  summarise(emissions = sum(emissions)) %>%
  ungroup()

# create the hierarchial dataset
parents_top <- tibble(
  ids = "<b>Greenhouse\nGas\nEmissions</b>",
  emissions = sum(ghg_tree$emissions),
  parents = "",
  labels = "<b>Greenhouse\nGas\nEmissions</b>")

parents <- ghg_tree %>%
  group_by(sg_categories) %>%
  summarise(emissions = sum(emissions, na.rm = TRUE)) %>%
  mutate(parents = "<b>Greenhouse\nGas\nEmissions</b>",
         labels = sg_categories,
         ids = paste0("<b>Greenhouse\nGas\nEmissions</b> - ", sg_categories)) %>%
  select(-sg_categories)

second_level <- ghg_tree %>%
  mutate(ids = paste0("<b>Greenhouse\nGas\nEmissions</b> - ", sg_categories, " - ", source_name),
         labels = source_name,
         parents = paste0("<b>Greenhouse\nGas\nEmissions</b>", " - ", sg_categories)) %>%
  select(-sg_categories) %>%
  group_by(ids, parents, labels) %>%
  summarise(emissions = sum(emissions, na.rm = TRUE))

third_level <- ghg_tree %>%
  drop_na(source_sub_cat) %>%
  mutate(ids = paste0("<b>Greenhouse\nGas\nEmissions</b> - ", sg_categories, " - ", source_name, " - ", source_sub_cat),
         parents = paste0("<b>Greenhouse\nGas\nEmissions</b> - ", sg_categories, " - ", source_name)) %>%
  rename(labels = source_sub_cat) %>%
  ungroup() %>%
  select(ids, parents, labels, emissions)

# combine all levels or heirarchy
ghg_hierarchy <- bind_rows(parents_top, parents, second_level, third_level)
```
Overview 
=======================================================================

Row
-----------------------------------------------------------------------
### `r emission_type` Emissions since 1990 to the latest available data
```{r warnings = FALSE, fig.width=15, fig.height = 9}

# Calculate total greenhouse gas emissions per sector per year
ghg_sector_total <- scot_ghg_clean %>%
  group_by(year, sg_categories) %>%
  summarise(total_emissions = sum(emissions)) %>%
  rename(sector = sg_categories) %>%
  mutate(sector = recode(sector, 
                                   "Domestic\nTransport" = "Domestic Transport",
                                   "Business\n&\nIndustry" = "Business & Industry")) 

# Order the sectors by greatest decrease in emissions
ghg_sector_total_plot <- ghg_sector_total %>%
  group_by(sector) %>%
  arrange(year) %>%
  mutate(emission_drop = last(total_emissions) - first(total_emissions),
         emission_drop = replace_na(emission_drop, 0)) %>%
   ungroup() %>%
  mutate(sector = fct_reorder(sector, emission_drop))

#colours to match sunburst
colors <- viridis(10, direction = -1, end = 0.9)

#create list of sectors with colors ordered by last year emissions
sector_list <- ghg_sector_total_plot %>%
  arrange(year) %>%
  filter(year == last_year) %>%
  select(sector, total_emissions, emission_drop) %>%
  arrange (desc(total_emissions))
#add colors to df to match sunburst
sector_list$sector_colors = colors

#arrange colours to match plot order
sector_list_color <- sector_list %>%
  arrange (emission_drop) %>%
  select (sector_colors)

# Label positionings for the stacked area plot
ghg_subs <- ghg_sector_total_plot %>%
  filter(year == 1995) %>%
  ungroup() %>%
  arrange(desc(sector)) %>%
  mutate(cum = cumsum(total_emissions),
         gap = cum - lag(cum),
         gap = coalesce(gap, 0.5),
         position = cum - (gap/2))

#stacked area plot
  ghg_sector_total_plot %>%
  ggplot() +
  aes(x = year, y = total_emissions, group = sector, fill = sector) +
  geom_area() +
  scale_fill_manual(values = as.matrix(sector_list_color))+
  theme_light() +
  scale_x_continuous(breaks = seq(first(ghg_sector_total_plot$year),
                                  last(ghg_sector_total_plot$year), 2)) +
    geom_text(data = ghg_subs,
          aes(x = 1995, y = position,
              label = sector),
          colour = "black",
          hjust = 0) +
  # labels
    theme(legend.position = "none",
         panel.grid = element_blank(),
         axis.text = element_text(face = "bold")) +
   labs(x = "Year",
        y = "Total Greenhouse Gas Emissions (Mt CO2)")
 
```


### Breakdown of `r last_year` Data - Click on a sector to expand the view
```{r warning = FALSE, message = FALSE, fig.height= 8, fig.width = 10}

# create sunburst
plot_ly(
  ghg_hierarchy,
  ids = ~ids,
  labels = ~ labels,
  parents = ~ parents,
  values = ~ emissions,
  type = 'sunburst',
  branchvalues = 'total',
  # orientation of text labels
  insidetextorientation='radial',
  hoverinfo = "text",
        hovertext = ~paste0("<b>",labels,"</b><br>", round(emissions, 2), " Mt")) %>%
  layout(colorway = colors)

```

Row {data-height=60}
-----------------------------------------------------------------------

### __Emissions by sector. The largest emissions reductions are furthest left and those that have increased furthest right__

Row
-----------------------------------------------------------------------
### [Energy Supply](https://scotland.shinyapps.io/sg-scottish-energy-statistics/?Section=WholeSystem&Chart=EnSupplyEmissions)

```{r}

simp_chart_plot("Energy Supply")

```

### Business & Industry

```{r}
# Show only "Business & Industry"

simp_chart_plot("Business & Industry")

```

### Agriculture

```{r}
# Show only Agriculture Emissions

simp_chart_plot("Agriculture")

```

### [Residential](../STAClimateDashboard/code/r/home_energy/home_energy_markdown/home_energy_plots.html)

```{r}

simp_chart_plot("Residential")

```

### [Transport](../STAClimateDashboard/code/r/cycling_and_walking/markdown/sta-climate-change_cycling.html)

```{r}
# Show only Transport Emissions

simp_chart_plot("Transport")


```

### Public

```{r}

simp_chart_plot("Public")

```

### Waste Management

```{r}

simp_chart_plot("Waste Management")

```

### Development

```{r}

simp_chart_plot("Development")

```

### Forestry

```{r}

simp_chart_plot("Forestry")

```

### International Aviation & Shipping

```{r}

simp_chart_plot("International Aviation and Shipping")

```

About
=======================================================================

```{=html}
<p>&nbsp;<br />
<!-- A summary sheet may be downloaded &ndash; <a href="cycle_counters_info_cards.pdf">cycle_counters_info_cards.pdf</a -->

<h3>About the Data</h3>

<p>
The data comes from an Officail Statistics Publication for Scotland. The full report is available via <a href="https://www.gov.scot/publications/scottish-greenhouse-gas-statistics-1990-2019/documents/">this link.</a> The report and data were published on the 15th June 2021 and cover the data up to the end of 2019.
</p>
<p>
Total greenhouse gasses are the sum total of the emissions from the following: CH, CO2, HFCs, N2O, NF3, PFCs, SF6. The unit used for emissions is megatonnes of CO2, or equivalent in terms of their global warming potential for other GHGs (as calculated by IPCC described further here. Only greenhouse gas emitters have been included in this visualisation - carbon sinks in land use and forestry are not displayed.
</p>

```